---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---

# Hypothesis:
There is a significant seasonal variation in indirect greenhouse gas emissions from household consumption in major Japanese cities.

## Install required packages if not installed
```{r}
if (!requireNamespace("tidyverse", quietly = TRUE)) install.packages("tidyverse")
if (!requireNamespace("readxl", quietly = TRUE)) install.packages("readxl")
if (!requireNamespace("car", quietly = TRUE)) install.packages("car")
if (!requireNamespace("agricolae", quietly = TRUE)) install.packages("agricolae")
```

## Load required libraries
```{r}
library(tidyverse)
library(tidymodels)
library(readxl)
library(car)
library(agricolae)
```

## Load the data from the Excel file
```{r}
# Set working directory
setwd("C:/Users/OMBATI/Desktop/r codes/dec-1.3")


file_path <- "12. gasoline_intensity.xlsx"
sheets <- c("2011", "2012", "2013", "2014", "2015")
```


## Combining the excel sheets
```{r}
# Create an empty list to store data frames
data_list <- list()

# Read data from each sheet and store in the list
for (sheet in sheets) {
  df <- read_excel(file_path, sheet = sheet)
  data_list[[sheet]] <- df
}

# Combine data frames into one
combined_data <- bind_rows(data_list, .id = "Year")

```

# EDA(Exploratory Data Analysis)
## Top rows of combined data
```{r}
head(combined_data)
```

## Bottom rows
```{r}
tail(combined_data)
```

## Statistical  summary
```{r}
summary(combined_data)
```
## Structure of the dataset
```{r}
str(combined_data)
```

## View of the columns of the dataset
```{r}
glimpse(combined_data)
```

## Convert the Year column into factor data type
```{r}
combined_data <- combined_data %>%
                 mutate_if(is.character,as.factor)
```





## Overview of the dataset
```{r}
combined_data %>% skimr::skim()
```


## Time Series Plot for a Specific Region(Tokyo)
```{r}
ggplot(combined_data, aes(x = as.factor(Month), y = Tokyo, group = Year, color = Year)) +
  geom_line() +
  labs(title = "Time Series Plot for Tokyo",
       x = "Month", y = "Value",
       color = "Year")

```

## Boxplot for a Specific Month
```{r}
ggplot(combined_data, aes(x = as.factor(Month))) +
  geom_boxplot() +
  labs(title = "Boxplot for Months",
       x = "Year", y = "Value")

```
## Scatterplot for Two Regions(Tokyo and Osaka)
```{r}
ggplot(combined_data, aes(x = Tokyo, y = Osaka, color = as.factor(Year))) +
  geom_point() +
  labs(title = "Scatterplot for Tokyo and  Osaka",
       x = "Tokyo", y = "Osaka", color = "Year")

```

## ANOVA for Seasonal Variation:
```{r}
model_anova <- aov(Sapporo ~ as.factor(Month), data = combined_data)
summary(model_anova)
```

## Linear Regression for Trend Over the Years:
```{r}
model_lm <- lm(Sapporo ~ as.numeric(Year), data = combined_data)
summary(model_lm)
```


##  Seasonal Decomposition of Time Series:
```{r}
decomposed_series <- decompose(ts(combined_data$Sapporo, frequency = 12))
plot(decomposed_series)
```

##  Paired t-test for Before and After a Certain Year:
```{r}
before_year <- subset(combined_data, Year %in% c(2011, 2012, 2013))
after_year <- subset(combined_data, Year %in% c(2014, 2015))
t_test_result <- t.test(before_year$Sapporo, after_year$Sapporo)
print(t_test_result)
```

